---
title: "US Murders"
author: "Spencer Varadi"
date: "8/9/2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12)
```
## Professor Plum with a Candle Stick in the Library
Homicide data from Kaggle was used to determine the mean annual murders per city reported from 1980 to 2014. This is not controlled by population. Data can be found here: https://www.kaggle.com/murderaccountability/homicide-reports

For an interactive experience using this data, please visit [my Homicides Shiny App](https://tghrsolutions.shinyapps.io/homicides/ "My app")

We will go through some exploratory analysis of

- Total murders over time
- The usual suspects
-  Weapons of choice

The analysis does not take population into consideration.

Lat and Lon provided using ggmap: D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2. The R Journal, 5(1), 144-161. URL http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf

## Load libraries and Data
```{r message = FALSE, echo = TRUE}
library(ggplot2)
library(plotly)
library(ggmap)
library(plyr)
library(reshape2)
library(scales)
homicides <- readRDS("data/homicides.rds")
# In case we want only a section of years
years <- c(1980:2015)
# Only interested in the US
countries <- c("USA", "United States")
homicides <- homicides[homicides$country %in% countries,]
```

## Total murders over time
```{r fig.align='center', echo = FALSE, message = FALSE}
freqYear <- count(homicides, vars = c("Year"))
filteredYear <- freqYear[freqYear$Year %in% years,]
p <- qplot(Year, freq, data = filteredYear) +
           geom_smooth() + labs(x = "Year", y = "Total Murders in US")
ggplotly(p)
```

<!-- ## Ussual Suspects -->
<!-- Proportion of murders commited each year by relationship to victim -->

<!-- Unknown subjects removed -->
<!-- ```{r fig.align='center', echo = FALSE, message = FALSE} -->
<!-- # relationship <- count(homicides, vars = c("Year","Relationship")) -->
<!-- relationship <- prop.table(table(homicides$Year[homicides$Relationship != "Unknown"],homicides$Relationship[homicides$Relationship != "Unknown"]), 1) -->
<!-- relationship <- as.data.frame(melt(relationship)) -->
<!-- names(relationship) <- c("Year","Relationship","Proportion") -->
<!-- relationship$percent <- percent(relationship$Proportion) -->
<!-- # relationship <- relationship[order(- relationship$Proportion),] -->
<!-- # aggregate(freq ~ Year + Relationship, relationship, FUN = function(x) x/sum(x)) -->
<!-- # dcast(relationship, Year ~ Relationship) -->

<!-- # realtionshipPer <- as.data.frame(lapply(dcast(relationship, Year ~ Relationship),function(x) x/sum(x))) -->
<!-- p <- ggplot() + geom_area(aes(y = Proportion, x = Year, fill = Relationship), stat = "identity", data = relationship -->
<!--           ) # + ggtitle("Proportion of murders commited each year by relationship to victim") -->
<!-- # summedByYear <- aggregate(freq ~ Year, filteredYear, sum) -->
<!-- ggplotly(p, hoverinfo = "text", text = ~paste('Year: ', Year)) -->

<!-- ``` -->

## Weapons of choice
```{r fig.align='center', echo = FALSE, message = FALSE}
weapons <- prop.table(table(homicides$Year,homicides$Weapon), 1)
weapons <- as.data.frame(melt(weapons))
names(weapons) <- c("Year","Weapon","Proportion")
weapons$percent <- percent(weapons$Proportion)
# aggregate(freq ~ Year + Relationship, relationship, FUN = function(x) x/sum(x))
# dcast(relationship, Year ~ Relationship)
# 
# realtionshipPer <- as.data.frame(lapply(dcast(relationship, Year ~ Relationship),function(x) x/sum(x)))
p <- ggplot() + geom_area(aes(y = Proportion, x = Year, fill = Weapon), stat = "identity", data = weapons) + ggtitle("Proportion of murders commited each year by weapon used")
# summedByYear <- aggregate(freq ~ Year, filteredYear, sum)
ggplotly(p, hoverinfo = "text", text = ~paste('Year: ', Year))
```

